---
title: "Simulation and summary of simulation results"
output: html_document
---

## `maicChecks` version 0.2.0 is required

To run the exact matching, R package `maicChecks` version 0.2.0 is required. The package can be downloaded from CRAN.

## Simulate data, perform matching, summarize results

Although each simulation takes less than 0.1 second machine time on a standard laptop, the user run time can be much longer. To run all 10,000 simulation, please expect up to 2 hours user run time.

The code below runs 10 simulations. To change this, set `rs_k <- rand_seeds`.

```{r warning=FALSE, message=FALSE}
## 'fn_simdt.R' defines the parameters for the simulation
## ... and the function for running 1 simulation
source('fn_simdt.R')
##
set.seed(20250105)
rand_seeds <- sample(1:100000, size = 10000, replace = FALSE)
## ::::::::::::::::::::::::::::::::: ##
rs_k <- rand_seeds[1:10] ## ten runs ##
## ::::::::::::::::::::::::::::::::: ##
t.start <- proc.time()
sim10k <- mapply(sim.dt.fn,
                 r_and_s = rs_k, 
                 n1_obs = 300, 
                 n2_obs = 300
)
t.end <- proc.time()
t.start
t.end
t.end - t.start ## total run time
rm(list = ls()[!ls() %in% c('rand_seeds', 'sim10k', 't.start', 't.end')])
save.image("sim10ksumm.RData")
```

## Summary 

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
require(ggplot2)
source('funs.R')
```

### Histograms of ESS

Figure 5 in the manuscript

```{r ess.hist, fig.width=10, fig.height=6, warning=FALSE, message=FALSE}
## combine all ess
selected_elements <- sim10k[2,]
ess_df <- do.call(rbind, selected_elements)
## prepare data for histogram
ess_long <- ess_df %>% 
  pivot_longer(cols = c("ess.ipd1.unc", "ess.ipd2.unc",
                        "ess.ipd1.con", "ess.ipd2.con",
                        "ess.ipd1.ps",  "ess.ipd2.ps"),
               names_to = 'type',
               values_to = 'ess') %>%
  mutate(method = ifelse(type %in% c('ess.ipd1.unc', 'ess.ipd2.unc'), 
                         'Unconstrained', 
                         ifelse(type %in% c('ess.ipd1.con', 'ess.ipd2.con') ,
                                'Constrained', 'PS weights')),
         method = factor(method, levels = c('Unconstrained', 'Constrained', 'PS weights')),
         study = ifelse(type %in% c('ess.ipd1.unc', 'ess.ipd1.con', "ess.ipd1.ps"),
                      'IPD A', 'IPD B')
         ) %>%
  select(study, ess, type, method) %>%
  filter(!is.na(ess))
##
p10k <- ggplot(data = ess_long, aes(ess)) +
  theme_bw() +
  geom_histogram(bins = 30,
                 aes(y = after_stat(count / sum(count))),
                 fill = 'grey', color = 'black') +
  facet_grid(method ~ study) + ##
  xlab('Effective Sample Size') +
  ylab('Density') +
  theme(strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 13),   
        axis.title.y = element_text(size = 13),  
        axis.text.x = element_text(size = 12),    
        axis.text.y = element_text(size = 12))
print(p10k)
##
# ggsave('histEss10knum.png', 
#        plot = p10k, 
#        width = 10, 
#        height = 6, 
#        units = 'in',
#        dpi = 300)
```

### Summary statistics of maximum (largest) weights

Table 3 in the manuscript

```{r summ.maxWts}
## find max standardized means
colns_with_weights <- c("wt.unc.sdz", "wt.con.sdz", "wt.ps.sdz")
max_weights <- lapply(sim10k[1,], 
                     get_max, ## get_max() is in funs.R
                     cols = colns_with_weights)
max_weights <- do.call(rbind, max_weights)
##
x <- summary(max_weights)
x <- t(x)
x <- as.data.frame(x)
x <- separate(x, col = Freq, into = c("summ", "stats"), sep = ":") %>%
  select(-Var2)
x_wide <- x %>% 
  pivot_wider(names_from = Var1, values_from = stats) %>%
  filter(!is.na(summ))
x.sd <- apply(max_weights, 2, sd, na.rm=TRUE)
x_summ <- data.frame(rbind(c('# of simulations', rep(10000, 3)),
                           x_wide, 
                           c('sd', round(x.sd, 3))))
print(x_summ)
```

### Summary of difference in response variable Y and Yc

Figure 6 in manuscript: boxplot of mean difference in response variable Yc

```{r}
selected_elements <- sim10k[5,]
ycbar_df <- do.call(rbind, selected_elements) 
## ybar_df includes also observed and weighted Y means
ycbar_long <- ycbar_df %>% 
  pivot_longer(cols = c("diff.obs", "diff.unc", "diff.con", "diff.ps"),
               names_to = 'type',
               values_to = 'ycbar.diff') %>%
  mutate(method = ifelse(type == 'diff.obs', 'Observed',
                         ifelse(type == 'diff.unc',  'Unconstrained',
                                ifelse(type == 'diff.con', 'Constrained', 'PS weighted'))
                        ),
         method = factor(method, 
                         levels = c('Observed', 'Unconstrained',
                                    'Constrained', 'PS weighted'))
         ) %>%
  select(ycbar.diff, type, method) %>%
  filter(!is.na(ycbar.diff))
## reverse the method levels for boxplot
ycbar_long <- ycbar_long %>% 
  mutate(method = factor(method, 
                         levels = c('PS weighted', 
                                    'Constrained',
                                    'Unconstrained', 
                                    'Observed'))
  )
```

```{r ycbarDiff, fig.width=10, fig.height=6}
p24k <- ggplot(ycbar_long, aes(ycbar.diff, method)) +
  geom_boxplot(orientation = 'y') + 
  geom_vline(xintercept = 0, 
             color = 'red', 
             linetype = 'dashed',
             lwd = 1) +
  theme_bw() +
  scale_x_continuous(breaks = seq(-2, 2.55, 0.5)) +
  theme(
    axis.title.x = element_text(size = 13),   
    axis.title.y = element_text(size = 13),  
    axis.text.x = element_text(size = 12),    
    axis.text.y = element_text(size = 12)     
  ) +
  xlab('Difference in mean responses') +
  ylab(NULL) 
##
print(p24k)
# ggsave('boxYcbar.png', 
#        plot = p24k, 
#        width = 10, 
#        height = 3, 
#        units = 'in',
#        dpi = 300)
```


Figure 7 in manuscript: boxplot of mean difference in response variable Y

```{r}
selected_elements <- sim10k[4,]
ybar_df <- do.call(rbind, selected_elements) 
## ybar_df includes also observed and weighted Y means
ybar_long <- ybar_df %>% 
  pivot_longer(cols = c("diff.obs", "diff.unc", "diff.con", "diff.ps"),
               names_to = 'type',
               values_to = 'ybar.diff') %>%
  mutate(method = ifelse(type == 'diff.obs', 'Observed',
                         ifelse(type == 'diff.unc',  'Unconstrained',
                                ifelse(type == 'diff.con', 'Constrained', 'PS weighted'))
                        ),
         method = factor(method, 
                         levels = c('Observed', 'Unconstrained',
                                    'Constrained', 'PS weighted'))
         ) %>%
  select(ybar.diff, type, method) %>%
  filter(!is.na(ybar.diff))
## reverse the method levels for boxplot
ybar_long <- ybar_long %>% 
  mutate(method = factor(method, 
                         levels = c('PS weighted', 
                                    'Constrained',
                                    'Unconstrained', 
                                    'Observed'))
  )
```

```{r ybarDiff, fig.width=10, fig.height=6}
p22k <- ggplot(ybar_long, aes(ybar.diff, method)) +
  geom_boxplot(orientation = 'y') + 
  geom_vline(xintercept = 0, 
             color = 'red', 
             linetype = 'dashed',
             lwd = 1) +
  theme_bw() +
  scale_x_continuous(breaks = seq(-2, 2.55, 0.5)) +
  theme(
    axis.title.x = element_text(size = 13),   
    axis.title.y = element_text(size = 13),  
    axis.text.x = element_text(size = 12),    
    axis.text.y = element_text(size = 12)     
  ) +
  xlab('Difference in mean responses') +
  ylab(NULL) 
##
print(p22k)
# ggsave('boxYbar.png', 
#        plot = p22k, 
#        width = 10, 
#        height = 3, 
#        units = 'in',
#        dpi = 300)
```

### Summary of run-time

```{r}
selected_elements <- sim10k[6,]
second_rows <- lapply(selected_elements, function(df) df[2, ])
# Combine these rows into a new dataframe
run_times <- do.call(rbind, second_rows) 
rownames(run_times) <- 1:nrow(run_times)
summary(run_times)
```

## Summary statistics of a signle simulated pair of IPDs

This is the 110th simulated data. 

### ESS

```{r}
s.num <- 110
simdt110_ess <- sim10k[2, s.num]$ess
print(simdt110_ess)
```

### Summary statistics of the covariates

Table 4 in the manuscript

```{r summ1}
simdt110_means <- sim10k[3, s.num]$all_means %>%
  select(-c('v.names', 'abs.smd')) ## work on abs.smd separately
## get abs.smd for later
x_smd <- sim10k[3, s.num]$all_means %>%
  select(v.names, abs.smd) %>%
  mutate_if(is.numeric, round, digits = 3) 
## save variable names
v_names <- sim10k[3, s.num]$all_means$v.names 
##
## :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::##
## manually fill in the reference levels for binary/categorical variables ##
## :::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::: ##
##
x <- data.frame(t(simdt110_means))
colnames(x) <- v_names
x <- x %>% mutate( ## will work on abs.smd later
  ## for binary variables, all references are 'missing'
  X1.A = 1 - X1.B,
  X2.A = 1 - X2.B,
  X6.A = 1 - X6.B,
  X7.A = 1 - X7.B,
  X11.A = 1 - X11.B,
  X12.A = 1 - X12.B,
  X13.A = 1 - X13.B,
  ## for categoricals, only unconstrained means are missing
  X3.A = ifelse(is.na(X3.A), 1 - X3.B - X3.C - X3.D, X3.A),
  X8.A = ifelse(is.na(X8.A), 1 - X8.B - X8.C, X8.A),
  X14.A = ifelse(is.na(X14.A), 1 - X14.B - X14.C - X14.D - X14.E, X14.A)
  ) %>%
  select(X1.A, X1.B, X2.A, X2.B, X3.A, X3.B, X3.C, X3.D,
         X4, X5, X6.A, X6.B, X7.A, X7.B, X8.A, X8.B, X8.C,
         X9, X10, X11.A, X11.B, X12.A, X12.B, X13.A, X13.B,
         X14.A, X14.B, X14.C, X14.D, X14.E, X15) %>%
  t(.) %>%
  data.frame(.) %>%
  mutate_if(is.numeric, round, digits = 3) 
##
## ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::: ##
## separating out categorical variables and convert to percentages ##
## ::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::::: ##
##
x_cat <- x %>%
  select_if(is.numeric) %>%
  filter(!(rownames(x) %in% c('X4', 'X5', 'X9', 'X10', 'X15'))) %>%
  mutate_if(is.numeric, ~ . * 100) %>%
  mutate(v.names = row.names(.)) %>%
  select('v.names', 'ipd1.obs', 'ipd2.obs', 
         'ipd12.unc', 'ipd12.con', 
         'ipd1.ps', 'ipd2.ps') 
##
## ::::::::::::::::::::::::::::::::::::::::::: ##
## combine x_cat with the continuous variables ##
## ::::::::::::::::::::::::::::::::::::::::::: ##
##
x_all <- x %>%
  filter(rownames(x) %in% 
           c('X4', 'X5', 'X9', 'X10', 'X15')) %>%
  mutate(v.names = row.names(.)) %>%
  select('v.names', 'ipd1.obs', 'ipd2.obs', 
         'ipd12.unc', 'ipd12.con', 
         'ipd1.ps', 'ipd2.ps') %>%
  rbind(x_cat) %>%
  left_join(x_smd, by = 'v.names') %>%
  mutate(v.names = factor(v.names, 
                          levels = c('X1.A', 'X1.B', 
                                     'X2.A', 'X2.B',
                                     'X3.A', 'X3.B', 'X3.C', 'X3.D',
                                     'X4', 'X5', 
                                     'X6.A', 'X6.B',
                                     'X7.A', 'X7.B', 
                                     'X8.A', 'X8.B', 'X8.C', 
                                     'X9', 'X10', 
                                     'X11.A', 'X11.B', 
                                     'X12.A', 'X12.B', 
                                     'X13.A', 'X13.B', 
                                     'X14.A', 'X14.B', 'X14.C', 'X14.D', 'X14.E', 
                                     'X15'))) %>%
  arrange(v.names)
print(x_all)
```

### Summary statistics of the response variable Y and Yc

```{r}
simdt110_ycbar <- round(sim10k[5, s.num]$ycbar, 3)
print(simdt110_ycbar)
simdt110_ybar <- round(sim10k[4, s.num]$ybar, 3)
print(simdt110_ybar)
```

## Save workspace

```{r}
rm(list = ls()[!ls() %in% 
                 c('random_seeds', 'sim10k',
                   't.start', 't.end',
                   'ess_df', 'ess_long', 
                   'max_weights',
                   'p10k', 'p22k', 'p24k', 
                   'simdt110_means', 'simdt110_ybar',
                   'simdt110_ycbar', 'x_all',
                   'ybar_df', 'ybar_long', 
                   'ycbar_df', 'ycbar_long')])
##save.image("sim10ksumm.RData")
```

```{r}
sessionInfo()
```

