---
title: "Simulation and summary of simulation results"
output: html_notebook
---

```{r}
## 'fn_simdt.R' defines the parameters for the simulation
## ... and the function for running 1 simulation
source('fn_simdt.R')
##
set.seed(20250105)
rand_seeds <- sample(1:100000, size = 10000, replace = FALSE)
rs_k <- rand_seeds[1:10] ## ten runs
t.start <- proc.time()
sim10k <- mapply(sim.dt.fn,
                 r_and_s = rs_k, 
                 n1_obs = 300, 
                 n2_obs = 300
)
t.end <- proc.time()
t.start
t.end
t.end - t.start ## total run time
rm(list = ls()[!ls() %in% c('rs_10k', 'sim10k', 't.start', 't.end')])
## 
save.image("sim10k.RData")
## 
```


## 2025-01-23

Response variable Y is also simulation. It depends on X1, X3, X8, X9, X11, and X15. The dependency on X9 and X15 are the underlying continuous version of these variables; on X1, X3, X8, and X11 are categorical.

## Background

The simulations are done on updated, local copy of `exmLP.2ipd()` and `exmWt.2ipd()`. The new functions are in "08_emmEx/simulatedExamples15/fnUpdates/local_exmLP.R" and "local_exmWt.R". 

Simulation parameters can be found in "08_emmEx/simulatedExamples15/fnUpdates/check_local_fns_new.R". 

To run the simulations:
"08_emmEx/simulatedExamples15/fnUpdates/checks.R". 


## Summary 

```{r setup, echo=FALSE, eval=FALSE}
require(dplyr, quietly = TRUE)
require(tidyverse, quietly = TRUE)
source('../../funs.R')
getwd()
## load data here if need
load("~/OneDrive - Novartis Pharma AG/Documents/bb/aa_NewStat/08_emmEx/simulatedExamples15/sim10k/YandCatX1X3X8X11/fnUpdatesSim10KcatX1X3X8X11.RData")
local_sim10k_cat <- local_sim10k
```

### Summary of difference in response variable 

This is when the weighted mean difference is 0.

```{r}
selected_elements <- local_sim10k_cat[4,]
ybar_df_cat <- do.call(rbind, selected_elements) 
## summary
x <- summary(ybar_df_cat[,9:12])
x <- t(x)
x <- as.data.frame(x)
x <- separate(x, col = Freq, into = c("summ", "stats"), sep = ":") %>%
  select(-Var2)
x_wide_cat <- x %>% 
  pivot_wider(names_from = Var1, values_from = stats) %>%
  filter(!is.na(summ))
x.sd <- apply(ybar_df_cat[,9:12], 2, sd, na.rm=TRUE)
x.sd
x_ybar_cat <- data.frame(rbind(c('# of simulations', rep(10000, 4)),
                           x_wide_cat, 
                           c('sd', round(x.sd, 3))))
x_ybar_cat
flextable::flextable(x_ybar_cat)

xtable::xtable(x_ybar_cat, 
               caption = "Summary statistics of the mean difference in the response variable Y between IPD 1 and IPD 2 in their observed values, exact matched weighted values (unconstrained and constrained), and propensity score weighted values.", 
               label = "tab:summybar_cat")
```

Histogram of the mean difference in ybar (not used in manuscript)

```{r}
ybar_long_cat <- ybar_df_cat %>% 
  pivot_longer(cols = c("diff.obs", "diff.unc", "diff.con", "diff.ps"),
               names_to = 'type',
               values_to = 'ybar.diff') %>%
  mutate(method = ifelse(type == 'diff.obs', 'Observed',
                         ifelse(type == 'diff.unc',  'Unconstrained',
                                ifelse(type == 'diff.con', 'Constrained', 'PS weighted'))
                        ),
         method = factor(method, levels = c('Observed', 'Unconstrained', 'Constrained', 'PS weighted'))
         ) %>%
  select(ybar.diff, type, method) %>%
  filter(!is.na(ybar.diff)) ## %>%
##  filter(ybar.diff <= 1.25)
##
p20k <- ggplot(ybar_long_cat, aes(ybar.diff)) +
  theme_bw() +
  geom_vline(xintercept = 0, color = 'red', lwd = 0.5, linetype = 'dashed') +
  geom_histogram(bins = 30,
                 aes(y = after_stat(count / sum(count))),
                 fill = 'grey', color = 'black') +
  facet_grid(method ~ .) + ##
  scale_x_continuous(breaks = seq(-1, 1, 0.25),
                     labels = seq(-1, 1, 0.25)) +
  xlab('Difference in mean responses') +
  ylab('Density') +
  theme(strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12)
        )
print(p20k)
##
getwd()
ggsave('histYbar10KwY.png', 
       plot = p20k, 
       width = 10, 
       height = 6, 
       units = 'in',
       dpi = 300)
```

Boxplot of mean difference: presented in the manuscript

```{r}
## reverse the method levels for boxplot
ybar_long_cat <- ybar_long_cat %>% 
  mutate(method = factor(method, 
                         levels = c('PS weighted', 
                                    'Constrained',
                                    'Unconstrained', 
                                    'Observed'))
  )
##x0 <- data.frame(t(x_ybar[5,2:5]))
##colnames(x0) <- 'ybar.diff'
##x0$method <- row.names(x0)
```

```{r}
p22k_cat <- ggplot(ybar_long_cat, aes(ybar.diff, method)) +
  geom_boxplot(orientation = 'y') + 
  geom_vline(xintercept = 0, 
             color = 'red', 
             linetype = 'dashed',
             lwd = 1) +
  theme_bw() +
  scale_x_continuous(breaks = seq(-2, 2.55, 0.5),
                     limits = c(-2, 2.55)
                     ##labels = seq(-1, 1, 0.25)
                     ) +
  theme(
    axis.title.x = element_text(size = 13),   
    axis.title.y = element_text(size = 13),  
    axis.text.x = element_text(size = 12),    
    axis.text.y = element_text(size = 12)     
  ) +
  xlab('Difference in mean responses') +
  ylab(NULL) ##+
##  geom_point(data = x0, aes(ybar.diff))
##
print(p22k_cat)
getwd()
ggsave('boxYbar10kcat.png', 
       plot = p22k_cat, 
       width = 10, 
       height = 3, 
       units = 'in',
       dpi = 300)
```

Scatter plot of the mean difference

```{r}
p25k <- ggplot(ybar_df_cat, aes(diff.unc, diff.ps)) +
  geom_point()
print(p25k)
```

### Summary of run-time

```{r}
selected_elements <- local_sim10k_cat[5,]
second_rows <- lapply(selected_elements, function(df) df[2, ])
# Combine these rows into a new dataframe
run_times <- do.call(rbind, second_rows) 
rownames(run_times) <- 1:10000
summary(run_times)
```

```{r eval=FALSE}
## removed everything except local_sim10k_cat
save.image("~/OneDrive - Novartis Pharma AG/Documents/bb/aa_NewStat/08_emmEx/simulatedExamples15/sim10k/YandCatX1X8/fnUpdatedSim10KcatX1X8.RData")
```

