---
title: "Simulated Illustrative Examples"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
##
knitr::opts_chunk$set(eval = FALSE)
##
require(dplyr)
require(tidyverse)
require(ggplot2)
require(maicChecks) ## version 0.2.0
```

## Data

### Simulate data

Simulate 3 examples, each example consisting 2 IPDs which are used for matching. There 80 observations in IPD 1, 120 observations in IPD 2. Each IPD has two variables, x1 and x2.

```{r simulateData}
## define function to simulate data :::::::::::::::::::: ##
s.dt <- function(n.xy, mean.x, mean.y, sd.x, sd.y, cor.xy){
  cov.xy <- cor.xy * sd.x * sd.y
  s.igma <- matrix(c(sd.x^2, cov.xy, cov.xy, sd.y^2),
                   nrow = 2, byrow = TRUE)
  ipd <- data.frame(mvtnorm::rmvnorm(n = n.xy,
                                     mean = c(mean.x, mean.y),
                                     sigma = s.igma))
  colnames(ipd) <- c('x1', 'x2')
  return(ipd)
}
## ::::::::::::::::::::: example 1 ::::::::::::::::::::: ##
set.seed(10)
dt1.ipd1 <- s.dt(n.xy = 80, 
                 mean.x = 0, 
                 mean.y = 0, 
                 sd.x = 1, 
                 sd.y = 2, 
                 cor.xy = 0.5)
set.seed(12)
dt1.ipd2 <- s.dt(n.xy = 120, 
                 mean.x = 0, 
                 mean.y = 0, 
                 sd.x = 1, 
                 sd.y = 2, 
                 cor.xy = 0.75)
## ::::::::::::::::::::: example 2 ::::::::::::::::::::: ##
dt2.ipd1 <- dt1.ipd1 
dt2.ipd2 <- dt1.ipd2
dt2.ipd2$x1 <- dt1.ipd2$x1 * (-1) + 1 
dt2.ipd2$x2 <- dt1.ipd2$x2 + 0.5
## ::::::::::::::::::::: example 3 ::::::::::::::::::::: ##
dt3.ipd1 <- dt1.ipd1
dt3.ipd2 <- dt1.ipd2
dt3.ipd2$x1 <- dt1.ipd2$x1 * (-1) + 1.5 * max(dt1.ipd1$x1) 
##
## putting them together ::::::::::::::::::::::::::::::: ##
##
dt.all <- data.frame(rbind(dt1.ipd1, dt1.ipd2, 
                           dt2.ipd1, dt2.ipd2, 
                           dt3.ipd1, dt3.ipd2
)) %>%
  mutate(example = c(rep('example 1', 
                         nrow(dt1.ipd1) + nrow(dt1.ipd2)),
                     rep('example 2', 
                         nrow(dt2.ipd1) + nrow(dt2.ipd2)),
                     rep('example 3', 
                         nrow(dt3.ipd1) + nrow(dt3.ipd2))),
         ipd = c(rep(1, nrow(dt1.ipd1)),
                 rep(2, nrow(dt1.ipd2)),
                 rep(1, nrow(dt2.ipd1)),
                 rep(2, nrow(dt2.ipd2)),
                 rep(1, nrow(dt3.ipd1)),
                 rep(2, nrow(dt3.ipd2))
         ))
```

### Exact matching 

Matching is performed on the three pairs of IPD. The R package `maicChecks` version 0.2.0 is required for the matching.   

```{r exactMatching}
## ::::::::::::::::::::: example 1 ::::::::::::::::::::: ##
## both constrained and unconstrained weights exist and are the same
## a. unconstrained
maicChecks::exmLP.2ipd(dt1.ipd1, 
                       dt1.ipd2, 
                       vars_to_match = c('x1', 'x2'),
                       mean.constrained = FALSE) 
os.dt1.f <- maicChecks::exmWt.2ipd(dt1.ipd1, 
                                   dt1.ipd2, 
                                   vars_to_match = c('x1', 'x2'),
                                   mean.constrained = FALSE)
## b. constrained
maicChecks::exmLP.2ipd(dt1.ipd1, 
                       dt1.ipd2, 
                       vars_to_match = c('x1', 'x2'),
                       mean.constrained = TRUE) 
os.dt1.t <- maicChecks::exmWt.2ipd(dt1.ipd1, 
                                   dt1.ipd2, 
                                   vars_to_match = c('x1', 'x2'),
                                   mean.constrained = TRUE)
## ::::::::::::::::::::: example 2 ::::::::::::::::::::: ##
## both constrained and unconstrained weights exist but are not the same
## a. unconstrained
maicChecks::exmLP.2ipd(dt2.ipd1, 
                       dt2.ipd2, 
                       vars_to_match = c('x1', 'x2'),
                       mean.constrained = FALSE) 
os.dt2.f <- maicChecks::exmWt.2ipd(dt2.ipd1, 
                                   dt2.ipd2, 
                                   vars_to_match = c('x1', 'x2'),
                                   mean.constrained = FALSE)
## b. constratined
maicChecks::exmLP.2ipd(dt2.ipd1, 
                       dt2.ipd2, 
                       vars_to_match = c('x1', 'x2'),
                       mean.constrained = TRUE) 
os.dt2.t <- maicChecks::exmWt.2ipd(dt2.ipd1, 
                                   dt2.ipd2, 
                                   vars_to_match = c('x1', 'x2'),
                                   mean.constrained = TRUE)
## ::::::::::::::::::::: example 3 ::::::::::::::::::::: ##
## unconstrained weights exist but constrained don't
## a. unconstrained
maicChecks::exmLP.2ipd(dt3.ipd1,
                       dt3.ipd2, 
                       vars_to_match = c('x1', 'x2'),
                       mean.constrained = FALSE) 
os.dt3.f <- maicChecks::exmWt.2ipd(dt3.ipd1, 
                                   dt3.ipd2, 
                                   vars_to_match = c('x1', 'x2'),
                                   mean.constrained = FALSE)
## b. constrained. the check below returns value 2
maicChecks::exmLP.2ipd(dt3.ipd1, dt3.ipd2, mean.constrained = TRUE) 
##
## ::::::::::::: adding the weights to dt.all :::::::::::::::: ##
##
dt.all <- dt.all %>%
  mutate(wts.uc = c(os.dt1.f$ipd1$exm.wts, os.dt1.f$ipd2$exm.wts,
                    os.dt2.f$ipd1$exm.wts, os.dt2.f$ipd2$exm.wts,
                    os.dt3.f$ipd1$exm.wts, os.dt3.f$ipd2$exm.wts),
         wts.c = c(os.dt1.t$ipd1$exm.wts, os.dt1.t$ipd2$exm.wts,
                   os.dt2.t$ipd1$exm.wts, os.dt2.t$ipd2$exm.wts,
                   rep(NA, nrow(dt3.ipd1) + nrow(dt3.ipd2)))
  )
```

### Observed means

Summarizing observed means. Also manually created x1 and x2 coordinates for labeling data points in the plots later. Weighted observed means of the two IPD are added later.

```{r meansObserved}
## ::::::::::::::::::::: example 1 :::::::::::::::::::::
means.orig.1 <- data.frame(x1 = c(mean(dt1.ipd1$x1), 
                                  mean(dt1.ipd2$x1)),
                           x2 = c(mean(dt1.ipd1$x2), 
                                  mean(dt1.ipd2$x2)),
                           ## x.txt and y.txt are for labeling the points on plots
                           x1.txt = c(mean(dt1.ipd1$x1) + 0.015, 
                                      mean(dt1.ipd2$x1) - 0.015),
                           x2.txt = c(mean(dt1.ipd1$x2), 
                                      mean(dt1.ipd2$x2))
)
## ::::::::::::::::::::: example 2 :::::::::::::::::::::
means.orig.2 <- data.frame(x1 = c(mean(dt2.ipd1$x1), mean(dt2.ipd2$x1)),
                           x2 = c(mean(dt2.ipd1$x2), mean(dt2.ipd2$x2)),
                           x1.txt = c(mean(dt2.ipd1$x1) + 0.15, 
                                      mean(dt2.ipd2$x1) - 0.08),
                           x2.txt = c(mean(dt2.ipd1$x2), 
                                      mean(dt2.ipd2$x2) - 0.08)
)
## ::::::::::::::::::::: example 3 :::::::::::::::::::::
means.orig.3 <- data.frame(x1 = c(mean(dt3.ipd1$x1), mean(dt3.ipd2$x1)),
                           x2 = c(mean(dt3.ipd1$x2), mean(dt3.ipd2$x2)),
                           x1.txt = c(mean(dt3.ipd1$x1) + 0.4, 
                                      mean(dt3.ipd2$x1) - 0.4),
                           x2.txt = c(mean(dt3.ipd1$x2), 
                                      mean(dt3.ipd2$x2))
)
## ::::::::::::::::::::: putting them together :::::::::::::::::::::
means.orig <- data.frame(rbind(means.orig.1, 
                               means.orig.2, 
                               means.orig.3
))
means.orig$example <- paste0('example ', rep(1:3, each = 2))
means.orig$which.means <- rep(c('IPD 1', 'IPD 2'), times = 3)
## below are dummy ipd codes for plotting
means.orig$ipd <- rep(c(31, 32), times = 3)
```

### Weighted means of unconstrained exact matching

Also manually create x1 and x2 coordiates used for labeling points in the plots later.

```{r meansUnconstrained}
## ::::::::::::::::::::: example 1 :::::::::::::::::::::
means.wtd.1.f <- data.frame(x1 = os.dt1.f$wtd.summ[3], 
                            x2 = os.dt1.f$wtd.summ[4],
                            ## x.txt and y.txt are for labelling the poitns on plots
                            x1.txt = os.dt1.f$wtd.summ[3], 
                            x2.txt = os.dt1.f$wtd.summ[4] + 0.006) 
## ::::::::::::::::::::: example 2 :::::::::::::::::::::
means.wtd.2.f <- data.frame(x1 = os.dt2.f$wtd.summ[3], 
                            x2 = os.dt2.f$wtd.summ[4],
                            x1.txt = os.dt2.f$wtd.summ[3], 
                            x2.txt = os.dt2.f$wtd.summ[4] - 0.08) 
## ::::::::::::::::::::: example 3 :::::::::::::::::::::
means.wtd.3.f <- data.frame(x1 = os.dt3.f$wtd.summ[3], 
                            x2 = os.dt3.f$wtd.summ[4],
                            x1.txt = os.dt3.f$wtd.summ[3], 
                            x2.txt = os.dt3.f$wtd.summ[4] - 0.2)
```

### Weight means of constrained exact matching

Also manually create x1 and x2 coordiates used for labeling points in the plots later.

```{r meansConstrained}
## ::::::::::::::::::::: example 1 ::::::::::::::::::::: ##
means.wtd.1.t <- data.frame(x1 = os.dt1.t$wtd.summ[3], 
                            x2 = os.dt1.t$wtd.summ[4],
                            ## x.txt and y.txt are for labelling the poitns on plots
                            x1.txt = os.dt1.f$wtd.summ[3], 
                            x2.txt = os.dt1.f$wtd.summ[4] - 0.004) 
## ::::::::::::::::::::: example 2 ::::::::::::::::::::: ##
means.wtd.2.t <- data.frame(x1 = os.dt2.t$wtd.summ[3], 
                            x2 = os.dt2.t$wtd.summ[4],
                            x1.txt = os.dt2.t$wtd.summ[3], 
                            x2.txt = os.dt2.t$wtd.summ[4] - 0.08) 
## :::::::: example 3: no constrained matching ::::::::: ##
means.wtd.3.t <- data.frame(x1 = NA, x2 = NA, x1.txt = NA, x2.txt = NA) 
```

### Combine observed means and weighted means

```{r meansAll}
## combine weighted means first :::::::::::::::::::::: ##
means.wtd <- data.frame(rbind(means.wtd.1.f, 
                              means.wtd.2.f, 
                              means.wtd.3.f,
                              means.wtd.1.t, 
                              means.wtd.2.t, 
                              means.wtd.3.t
))
means.wtd$example <- paste0('example ', rep(1:3, 2))
means.wtd$which.means <- rep(c('Unconstrained', 'Constrained'), each = 3)
## below are dummy ipd codes for plotting
means.wtd$ipd <- rep(c(33, 34), each = 3)
## putting together all the means :::::::::::::::::: ##
dt.means <- merge(means.orig, means.wtd, all = TRUE) %>% 
  arrange(example, ipd)
## ::::::::::::::::::::::::::::::::::::::::::::::::: ##
## coordinates for the four corners of the red box ...
## ... around the means in figure 2 of the paper
dt.means <- dt.means %>% group_by(example) %>%
  mutate(minx1 = min(x1, na.rm = TRUE) - 0.25,
         maxx1 = max(x1, na.rm = TRUE) + 0.25,
         minx2 = min(x2, na.rm = TRUE) - 0.35,
         maxx2 = max(x2, na.rm = TRUE) + 0.35
  ) %>%
  ungroup() %>%
  mutate(minx1 = ifelse(example == 'example 1',
                        minx1 + 0.1, minx1),
         maxx1 = ifelse(example == 'example 1',
                        maxx1 - 0.1, maxx1),
         minx1 = ifelse(example == 'example 3',
                        minx1 - 0.1, minx1),
         maxx1 = ifelse(example == 'example 3',
                        maxx1 + 0.1, maxx1)
  )
```

### Add weighted average of two observed IPD means 

Weighted observed means are added per first round of reviewer comment received in November 2024.

```{r meansWtdObserved}
n_ipd1 <- nrow(dt1.ipd1)
n_ipd2 <- nrow(dt1.ipd2)
##
dt.means.rev <- data.frame(matrix(NA, 
                                  nrow = 3, 
                                  ncol = ncol(dt.means)))
colnames(dt.means.rev) <- colnames(dt.means)
dt.means.rev <- dt.means.rev %>% 
  mutate(which.means = 'IPD 1 + IPD 2',
         example = paste0('example ', 1:3),
         ipd = max(dt.means$ipd) + 1, 
         x1 = dt.means[which(dt.means$which.means == 'IPD 1'),]$x1 *
           n_ipd1 / (n_ipd1 + n_ipd2) +
           dt.means[
             which(dt.means$which.means == 'IPD 2'),
           ]$x1 * n_ipd2 / (n_ipd1 + n_ipd2),
         x2 = dt.means[
           which(dt.means$which.means == 'IPD 1'),]$x2 * 
           n_ipd1 / (n_ipd1 + n_ipd2) +
           dt.means[which(dt.means$which.means == 'IPD 2'),]$x2 * 
           n_ipd2 / (n_ipd1 + n_ipd2),
         x1.txt = x1,
         x2.txt = ifelse(example == 'example 1', x2 + 0.005,
                         ifelse(example == 'example 2', 
                                x2 - 0.08,
                                x2 + 0.2)),
         minx1 = dt.means[which(dt.means$which.means == 'IPD 1'),]$minx1,
         maxx1 = dt.means[which(dt.means$which.means == 'IPD 1'),]$maxx1,
         minx2 = dt.means[which(dt.means$which.means == 'IPD 1'),]$minx2,
         maxx2 = dt.means[which(dt.means$which.means == 'IPD 1'),]$maxx2
  )
##
dt.means.rev <- data.frame(rbind(dt.means, dt.means.rev)) 
## in the manuscript they are called IPD A and IPD B
dt.means.rev <- dt.means.rev %>%
  mutate(which.means.o = which.means) %>%
  mutate(which.means = ifelse(which.means.o == 'IPD 1', 'IPD A',
                              ifelse(which.means.o == 'IPD 2', 'IPD B',
                                     ifelse(which.means.o == 'IPD 1 + IPD 2', 'IPD A + IPD B',
                                            which.means.o)))
  )
```

## Figures in the manuscript

### Figure 1

Three pairs of simulated IPD's.

```{r fig1, fig.height=2.67, fig.width=8}
fig.1 <- ggplot(dt.all, 
                aes(x1, x2, shape = factor(ipd))) +
  theme_bw() +
  geom_point() +
  scale_shape_manual(values = c(3, 1), 
                     labels = c('IPD A', 'IPD B')) +
  facet_wrap(. ~ example, scales = "free") +
  theme(legend.position = 'none', 
        legend.title = element_blank(),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.background=element_rect(fill = 0),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
##
fig.1
ggsave(plot = fig.1, file = 'fig1.jpeg', 
       width = 8, height = 2.67, units = 'in', dpi = 300)
```

### Figure 2 top row

Observed and weighted means with observed individual patient data

```{r fig2top, fig.height=2.67, fig.width=8}
fig.2top <- ggplot(data = dt.all, 
                   aes(x1, x2, 
                       color = factor(ipd), 
                       shape = factor(ipd) 
                   )) +
  geom_point(aes(size = wts.uc)) +
  geom_point(data = subset(dt.means.rev, !is.na(x1) & !is.na(x2)),
             aes(x1, x2, color = factor(ipd)), 
             size = 3
  ) + 
  theme_bw() +
  scale_shape_manual(values = c(3, 1, 18, 18, 15, 17, 19),
                     labels = c('IPD A', 'IPD B', '', '', '')) +
  scale_color_manual(values = c('grey65', 'grey65', 'black', 'black',
                                'red', 'blue', 'black'),
                     labels = c('IPD A', 'IPD B', '', '', '', '')) +
  geom_segment(data = dt.means.rev,
               aes(x = minx1, xend = maxx1,
                   y = minx2, yend = minx2),
               color = 'red') +
  geom_segment(data = dt.means.rev,
               aes(x = minx1, xend = maxx1,
                   y = maxx2, yend = maxx2),
               color = 'red') +
  geom_segment(data = dt.means.rev,
               aes(x = minx1, xend = minx1,
                   y = minx2, yend = maxx2),
               color = 'red') +
  geom_segment(data = dt.means.rev,
               aes(x = maxx1, xend = maxx1,
                   y = minx2, yend = maxx2),
               color = 'red') +
  facet_wrap(. ~ example, scales = "free") +
  theme(legend.position = 'none',
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 12),
        strip.background=element_rect(fill = 0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
fig.2top
ggsave(plot = fig.2top, file = 'fig2top.jpeg', 
       width = 8, height = 2.67, units = 'in', dpi = 300)
```

### Figure 2 bottom row

Observed and weighted means in the red box

```{r fig2bott, fig.height=2.67, fig.width=8}
fig.2bott <- dt.means.rev %>% 
  group_by(example) %>%
  ## use a difference scale for min/max as in fig.2top
  mutate(minx1 = min(x1, na.rm = TRUE),
         maxx1 = max(x1, na.rm = TRUE),
         minx2 = min(x2, na.rm = TRUE),
         maxx2 = max(x2, na.rm = TRUE)
  ) %>%
  distinct(example, minx1, maxx1, minx2, maxx2) %>%
  left_join(dt.all, by = 'example') %>%
  filter(x1 >= minx1 & x1 <= maxx1 & x2 >= minx2 & x2 <= maxx2) %>%
  ggplot(data = ., 
         aes(x1, x2,
             color = factor(ipd), 
             shape = factor(ipd))) +
  geom_point(aes(size = wts.uc)) +
  geom_point(data = subset(dt.means.rev, !is.na(x1) & !is.na(x2)),
             aes(x1, x2, color = factor(ipd)),
             size = 3) +
  theme_bw() +
  scale_shape_manual(values = c(3, 1, 18, 18, 15, 17, 19),
                     labels = c('IPD A', 'IPD B', '', '', '')) +
  scale_color_manual(values = c('grey65', 'grey65', 'black', 'black', 
                                'red', 'blue', 'black'),
                     ##c(4, 2, 1, 1, 1, 1),
                     labels = c('IPD A', 'IPD B', '', '', '')) +
  geom_text(data = subset(dt.means.rev, !is.na(x1) & !is.na(x2)), 
            aes(x1.txt, x2.txt, label = which.means)
  ) +
  facet_wrap(. ~ example, scales = "free") +
  theme(legend.position = 'none', 
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 12),
        strip.background=element_rect(fill = 0)
  ) 
##
fig.2bott
ggsave(plot = fig.2bott, file = 'fig2bott.jpeg', 
       width = 8, height = 2.67, units = 'in', dpi = 300)
```


```{r cleanup}
rm(list = setdiff(ls(), 
                  c('dt.all', 'dt.means.rev', 'fig.1', 'fig.2top', 'fig.2bott')))
save.image('illustrative.RData')
```

